---
title: "input-output-basics"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{input-output-basics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(inputoutput)
```

## Common variables

Let `X` be the input-output matrix and `d` the total final demand vector.

```{r}
X <- transaction_matrix
d <- wage_demand_matrix[, "final_total_demand"]
w <- wage_demand_matrix[, "wage"]
```

## Leontief approach

### Input requirement matrix

Let `A` be the direct coefficients matrix.

```{r}
A <- input_requirement(X, d)
```

### Leontief inverse matrix

```{r}
L <- leontief_inverse(A)
```

### Output multiplier

```{r}
output_multiplier(L)
```

### Household income multiplier

```{r}
h <- household_income_multiplier(L,w)
```

### Income multiplier

```{r}
i <- income_multiplier(L,w)
```

### Backward linkage

```{r}
bl <- backward_linkage(L)
bl_cv <- backward_linkage_cv(L)
```

### Forward linkage

```{r}
bl <- forward_linkage(L)
bl_cv <- forward_linkage_cv(L)
```

## Ghosh approach

### Output allocation matrix

Let `B` be the output allocation matrix.

```{r}
B <- output_allocation(X, d)
```

### Ghosh inverse matrix

```{r}
G <- ghosh_inverse(A)
```

### Input multiplier

```{r}
input_multiplier(G)
```
